---
title: "Model_4_All_models_without_performer"
author: "Srivastav Budugutta"
date: "2023-03-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Overview
>In this we will be investigating on the performers data.
Different types of performers are: individual names, performers with "featuring in between performers, & is there between performers, "+" is there when the artisits combine with other performers.


#initiating library

```{r}
rm(list = ls())

library(tidyverse)
```
## Function definitions
```{r}
clean_rating = function(dat){
  dat = dat %>% mutate(rating = if_else(rating<0,0,rating))
  
  return(dat)
}

change_track_explicit_to_numeric = function(dat){
  dat = dat %>% mutate(track_explicit = ifelse(track_explicit == FALSE,0,1))
  return(dat)
}

create_interactions_Variable = function(dat){
  dat = dat %>% mutate(interactive = danceability*energy)
  return(dat)
}

create_song_intensity = function(dat){
  dat = dat %>% mutate(song_intensity = danceability * energy * (10^(loudness/20)))
  return(dat)
}

create_acoustic_energy = function(dat){
  dat = dat %>% mutate(acoustic_energy = (1 - instrumentalness) * energy)
  return(dat)
}

create_instrumental_energy = function(dat){
  dat = dat %>% mutate(instrumental_energy = energy - (energy*(1-instrumentalness)))
  return(dat)
}

create_live_energy = function(dat){
  dat = dat %>% mutate(liveenergy = liveness*energy)
  return(dat)
}

# not working
create_danceability_variance = function(dat){
  dat = dat %>% mutate(danceability_variance = sd(danceability))
  return(dat)
}

# not working
create_temo_variance = function(dat){
  dat = dat %>% mutate(tempo_variance = sd(tempo))
  return(dat)
}



```

# Train Data Set
```{r}
songs = read_csv("data/analysisData.csv")


dim(songs)
songs = clean_rating(songs)
songs = change_track_explicit_to_numeric(songs)
songs %>% group_by(performer) %>% summarise(mean_rating = mean(rating))


# songs$track_explicit
head(songs)
min(songs$rating)
# geners_modify = songs %>% mutate(genre = str_extract_all(genre, regex("\'(.*?)\'", ignore_case = T)))
# geners_modify$genre

# we are trying to separate all the genre types
geners_modify = songs %>% separate_rows(genre, sep = ",\\s*")
geners_modify = geners_modify %>%  mutate(genre = str_replace_all(genre, "\\[|\\]|'|\"", ""))
geners_modify %>% group_by(genre) %>% count()
genre_mean_rating = geners_modify %>% group_by(genre) %>% summarize(avg_rating = mean(rating))
genre_mean_rating %>% filter(genre == "acoustic punk")
genre_mean_rating[geners_modify$genre,]
head(geners_modify)
geners_modify = geners_modify %>% mutate(track_duration = track_duration/(60*1000))

(geners_modify %>% group_by(genre) %>% summarize(avg_rating = mean(rating)) %>% arrange(avg_rating))
# geners_modify_1= geners_modify %>% group_by(genre) %>% summarize(avg_rating = mean(rating)) %>% ungroup() %>% mutate(genre = avg_rating)
# now we know what are the mean ratings for each genre now we will combine it to the song id.
geners_modify_1= geners_modify %>% group_by(genre) %>% mutate(avg_rating = mean(rating)) %>% ungroup() %>% mutate(genre = avg_rating)
geners_modify_1 = geners_modify_1 %>% group_by(id) %>% mutate(genre = mean(genre))
geners_modify_1 = geners_modify_1 %>%  select(-avg_rating)
mean(geners_modify_1$genre[c(1:4)])
geners_modify_1 %>% group_by(id)
geners_modify_2= distinct(geners_modify_1, .keep_all = T)

geners_modify_2 = create_interactions_Variable(geners_modify_2)
# geners_modify_2 = create_danceability_variance(geners_modify_2)
# geners_modify_2 = create_temo_variance(geners_modify_2)
geners_modify_2 = create_song_intensity(geners_modify_2)
geners_modify_2 = create_acoustic_energy(geners_modify_2)
geners_modify_2 = create_live_energy(geners_modify_2)
geners_modify_2 = create_instrumental_energy(geners_modify_2)
geners_modify_2
```


## Test





```{r}
songs_test = read_csv("data/scoringData.csv")
songs_test = change_track_explicit_to_numeric(songs_test)
songs_test = songs_test %>% separate_rows(genre, sep = ",\\s*")
songs_test = songs_test %>%  mutate(genre = str_replace_all(genre, "\\[|\\]|'|\"", ""))
songs_test = songs_test %>% mutate(track_duration = track_duration/(60*1000))

# songs_test_without_rating = songs_test
genre_mean_rating 
songs_test

# copying the corresponding values of the genre type to songs
index_value = match(songs_test$genre, genre_mean_rating$genre)
# index_value
songs_test$genre = genre_mean_rating$avg_rating[index_value] 
songs_test_1 = songs_test %>% group_by(id) %>% mutate(genre = mean(genre))
songs_test_1
songs_test_2= distinct(songs_test_1, .keep_all = T)
songs_test_2
songs_test_2 = create_interactions_Variable(songs_test_2)
# geners_modify_2 = create_danceability_variance(geners_modify_2)
# geners_modify_2 = create_temo_variance(geners_modify_2)
songs_test_2 = create_song_intensity(songs_test_2)
songs_test_2 = create_acoustic_energy(songs_test_2)
songs_test_2 = create_live_energy(songs_test_2)
songs_test_2 = create_instrumental_energy(songs_test_2)
dim(songs_test_2)

# merged_song_set = merge(songs_test, genre_mean_rating, by.x = "genre", by.y = "genre",all.x = T)
# merged_song_set
# songs_test$genre = merged_song_set$avg_rating
# songs_test
# songs_test_without_rating
#lets verify if the values for that particular song is given right average rating value.
# songs_test = songs_test %>% group_by(genre) %>% mutate(avg_rating = mean(rating)) %>% ungroup() %>% mutate(genre = avg_rating)
geners_modify_2


```



# check for best subset

```{r}
library(rpart); library(rpart.plot)
geners_modify_2[,4:length(geners_modify_2)]
tree = rpart(rating~.,data=geners_modify_2[,4:length(geners_modify_2)])
pred = predict(tree,newdata=songs_test_2)
# pred
# rmse_tree = sqrt(mean((pred-songs_test_2$rating)^2)); rmse_tree

# Tree with tuning

library(caret)
trControl = trainControl(method='cv',number = 5)
tuneGrid = expand.grid(.cp = seq(from = 0.001,to = 0.1,by = 0.001))
set.seed(617)
cvModel = train(rating~.,
                data=geners_modify_2[,4:length(geners_modify_2)],
                method="rpart",
                trControl = trControl,
                tuneGrid = tuneGrid)

cvModel$results

ggplot(data=cvModel$results, aes(x=cp, y=RMSE))+
  geom_line(size=0.5,alpha=0.2)+
  geom_point(color='brown')+
  theme_bw()+
  ggtitle(label=paste('Lowest RMSE is at a cp of ',cvModel$bestTune$cp))

rmse_cvTree = 14.61838

# Essemble MOdels


# Bagging
# Tuned Random Forest

trControl=trainControl(method="cv",number=5)
tuneGrid = expand.grid(mtry=1:4)
cvModel = train(rating~.,data=geners_modify_2[,4:length(geners_modify_2)],
                 method="rf",ntree=1000,trControl=trControl,tuneGrid=tuneGrid )
cvModel

```


# prediction

```{r}
#good subset:
# genre
# danceability
# track_duration
# speechiness
# interactive
# song_intensity
# loudness
# time_signature
# instrumental_energy
# liveenergy
# track_explicit

model_1 = lm(rating~genre+danceability+track_duration+speechiness+interactive+song_intensity+loudness+time_signature+instrumental_energy+liveenergy+track_explicit,geners_modify_2)
model_1_predict = predict(model_1,songs_test_2)
# length(songs_test_2$id)
# length(model_1_predict)
# songs_test_2[2812,]

# songs_test_2[,6]



```


# Saving the model

```{r}

submissionFile = data.frame(id = songs_test_2$id, rating = model_1_predict)
submissionFile[is.na(submissionFile)] = 24
write.csv(submissionFile, 'Model_4_All_models_without_performer.csv',row.names = F)
# head(songs_test)
# head(geners_modify)
# dim(geners_modify)


```




```{r}



```




```{r}



```