---
title: "Model_4_WW15_2_xgboost"
author: "Srivastav Budugutta"
date: "2023-03-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Overview
>In this we will be investigating on the performers data.
Different types of performers are: individual names, performers with "featuring in between performers, & is there between performers, "+" is there when the artisits combine with other performers.


#initiating library

```{r}
rm(list = ls())
library(tidyverse)
library(vtreat)
library(mice)
library(xgboost); library(caret)
library(randomForest)
library(ISLR2)
```
## Function definitions
```{r}
set.seed(1031)


clean_rating = function(dat){
  dat = dat %>% mutate(rating = if_else(rating<0,0,rating))
  
  return(dat)
}

change_track_explicit_to_numeric = function(dat){
  dat = dat %>% mutate(track_explicit = ifelse(track_explicit == FALSE,0,1))
  return(dat)
}

create_interactions_Variable = function(dat){
  dat = dat %>% mutate(interactive = danceability*energy)
  return(dat)
}

create_song_intensity = function(dat){
  dat = dat %>% mutate(song_intensity = danceability * energy * (10^(loudness/20)))
  return(dat)
}

create_acoustic_energy = function(dat){
  dat = dat %>% mutate(acoustic_energy = (1 - instrumentalness) * energy)
  return(dat)
}

create_instrumental_energy = function(dat){
  dat = dat %>% mutate(instrumental_energy = energy - (energy*(1-instrumentalness)))
  return(dat)
}

create_live_energy = function(dat){
  dat = dat %>% mutate(liveenergy = liveness*energy)
  return(dat)
}

# not working
create_danceability_variance = function(dat){
  dat = dat %>% mutate(danceability_variance = sd(danceability))
  return(dat)
}

# not working
create_temo_variance = function(dat){
  dat = dat %>% mutate(tempo_variance = sd(tempo))
  return(dat)
}

modify_performer_gener = function(dat){
  dat = dat %>% mutate(ph = ifelse(grepl("pinoy hip hop", x= dat$genre, ignore.case = T), "YES", "NO")) %>% mutate(stp = ifelse(grepl("stomp pop", x= dat$genre, ignore.case = T) |
                                                                                                                              grepl("portland hip hop", x= dat$genre, ignore.case = T) |
                                                                                                                              grepl("minnesota hip hop", x= dat$genre, ignore.case = T) |
                                                                                                                              grepl("k-hop", x= dat$genre, ignore.case = T), "YES", "NO")) %>% mutate(aq = ifelse(grepl("alt z", x= dat$genre, ignore.case = T)|
                         grepl("queer country", x= dat$genre, ignore.case = T), "YES", "NO")) %>%
  mutate(im = ifelse(grepl("italo dance", x= dat$genre, ignore.case = T)|
                         grepl("middle earth", x= dat$genre, ignore.case = T), "YES", "NO")) %>%
  mutate(cd = ifelse(grepl("corrido", x= dat$genre, ignore.case = T)|
                         grepl("deep regional mexican", x= dat$genre, ignore.case = T), "YES", "NO")) %>% mutate(ua = ifelse(grepl("uk funky", x= dat$genre, ignore.case = T)|
                         grepl("anarcho-punk", x= dat$genre, ignore.case = T), "YES", "NO")) %>% mutate(ua = ifelse(grepl("viral trap", x= dat$genre, ignore.case = T)|
                         grepl("theme", x= dat$genre, ignore.case = T), "YES", "NO")) %>% mutate(bc = ifelse(grepl("bassline", x= dat$genre, ignore.case = T)|
                         grepl("cartoon", x= dat$genre, ignore.case = T), "YES", "NO")) %>% mutate(bmv = ifelse(grepl("barbadian pop", x= dat$genre, ignore.case = T)|
                         grepl("melbourne bounce international", x= dat$genre, ignore.case = T) | grepl("vapor twitch", x= dat$genre, ignore.case = T), "YES", "NO")) %>% 
  mutate(ecggs = ifelse(grepl("etherpop", x= dat$genre, ignore.case = T)|
                         grepl("celtic", x= dat$genre, ignore.case = T) | grepl("gauze pop", x= dat$genre, ignore.case = T) | grepl("german soundtrack", x= dat$genre, ignore.case = T) | grepl("shiver pop", x= dat$genre, ignore.case = T), "YES", "NO")) %>% 
  
  mutate(dds = ifelse(grepl("dfw rap", x= dat$genre, ignore.case = T)|
                         grepl("danish pop", x= dat$genre, ignore.case = T) | grepl("steampunk", x= dat$genre, ignore.case = T), "YES", "NO")) %>%
  
  mutate(rd = ifelse(grepl("regional mexican pop", x= dat$genre, ignore.case = T)|
                         grepl("dubstep", x= dat$genre, ignore.case = T) | grepl("steampunk", x= dat$genre, ignore.case = T), "YES", "NO")) %>% 
  
  
  mutate(s = ifelse(grepl("slap house", x= dat$genre, ignore.case = T), "YES", "NO")) %>%
  
  mutate(sddfn = ifelse(grepl("swedish hard rock", x= dat$genre, ignore.case = T)|
                         grepl("deep norteno", x= dat$genre, ignore.case = T) | grepl("derby indie", x= dat$genre, ignore.case = T) | grepl("filter house", x= dat$genre, ignore.case = T) | grepl("norteno-sax", x= dat$genre, ignore.case = T), "YES", "NO")) %>% 
  
  mutate(an_1 = ifelse(grepl("austropop", x= dat$genre, ignore.case = T)|
                         grepl("neue deutsche welle", x= dat$genre, ignore.case = T), "YES", "NO")) %>%
  
  
  mutate(iss_1 = ifelse(grepl("indie singer-songwriter", x= dat$genre, ignore.case = T), "YES", "NO")) %>%
  
  mutate(cp_1 = ifelse(grepl("canadian latin", x= dat$genre, ignore.case = T)|
                         grepl("progressive electro house", x= dat$genre, ignore.case = T), "YES", "NO")) %>%  
  
  
  mutate(new_age = ifelse(grepl("new age", x= dat$genre, ignore.case = T), "YES", "NO")) %>%  
  
  mutate(gbl_1 = ifelse(grepl("german pop", x= dat$genre, ignore.case = T) |
                            grepl("breakbeat", x= dat$genre, ignore.case = T)|
                            grepl("liquid funk", x= dat$genre, ignore.case = T), "YES", "NO")) %>% 
  
  mutate(scandipop = ifelse(grepl("scandipop", x= dat$genre, ignore.case = T) , "YES", "NO")) %>%
  
  mutate(kaijs_1 = ifelse(grepl("k-pop boy group", x= dat$genre, ignore.case = T) |
                            grepl("alabama rap", x= dat$genre, ignore.case = T)|
                            grepl("italo house", x= dat$genre, ignore.case = T) |
                            grepl("jazz trio", x= dat$genre, ignore.case = T)|
                            grepl("ska revival", x= dat$genre, ignore.case = T), "YES", "NO")) %>%
  
   mutate(the_performer = ifelse(grepl("The", x = dat$performer, ignore.case = TRUE), "YES", "NO")) %>%  
  mutate(the_performer = ifelse(grepl("The", x = dat$performer, ignore.case = TRUE), "YES", "NO")) %>%
  
  mutate(group_performer = ifelse(grepl("\\s*&\\s*", x = dat$performer, ignore.case = TRUE) | 
                                    grepl("\\s*featuring\\s*", x = dat$performer, ignore.case = TRUE) |
                                    grepl("\\s+or\\s+", x = dat$performer, ignore.case = TRUE) |
                                    grepl("\\s*with\\s*", x = dat$performer, ignore.case = TRUE), "YES", "NO"))
  
  return(dat)
}

change_track_duration = function(dat){
  dat = dat %>% mutate(track_duration = track_duration/(60*1000))
}

```

# Reading prediction

```{r}

predictions = read_csv("data/predictions_gener_mean.csv")
head(predictions)
predictions = predictions %>% select(-Column1_orig)
colnames(predictions)
new_predictions = predictions %>% group_by(id_orig) %>% mutate(rating_predicted = mean(rating_predicted)) %>% select(-genre_orig)
new_predictions = distinct(new_predictions, .keep_all = T)
unique(new_predictions)
new_predictions
submissionFile = data.frame(id = new_predictions$id_orig, rating = new_predictions$rating_predicted)
write.csv(submissionFile, 'Azure_model_genre_mean.csv',row.names = F)

```

# Train Data Set
```{r}
songs = read_csv("data/analysisData.csv")
q4_sample_Data = createDataPartition(y = songs$rating, p=0.8, groups = 10, list = F)
songs_test = songs[-q4_sample_Data,]
songs_train = songs[q4_sample_Data,]
songs
songs_test

dim(songs)
songs_train = clean_rating(songs_train)
songs_test = clean_rating(songs_test)
songs_train = change_track_explicit_to_numeric(songs_train)
songs_test = change_track_explicit_to_numeric(songs_test) 
# songs %>% group_by(performer) %>% summarise(mean_rating = mean(rating))
# songs %>% summarize(mean_of_rating = mean(rating))

# songs$track_explicit
# head(songs)
# min(songs$rating)
# geners_modify = songs %>% mutate(genre = str_extract_all(genre, regex("\'(.*?)\'", ignore_case = T)))
# geners_modify$genre

# we are trying to separate all the genre types
# geners_modify = songs %>% separate_rows(genre, sep = ",\\s*")
# geners_modify = geners_modify %>%  mutate(genre = str_replace_all(genre, "\\[|\\]|'|\"", ""))
# geners_modify %>% group_by(genre) %>% count()
# geners_modify %>% filter(genre == "pinoy hip hop")
# genre_mean_rating = geners_modify %>% group_by(genre) %>% summarize(avg_rating = mean(rating))
# genre_mean_rating %>% filter(genre == "acoustic punk")
# order(genre_mean_rating)
# genre_mean_rating %>% arrange(desc(avg_rating)) %>% filter(avg_rating >50)
# head(geners_modify)
# geners_modify = geners_modify %>% mutate(track_duration = track_duration/(60*1000))
# 
# (geners_modify %>% group_by(genre) %>% summarize(avg_rating = mean(rating)) %>% arrange(avg_rating))
# # geners_modify_1= geners_modify %>% group_by(genre) %>% summarize(avg_rating = mean(rating)) %>% ungroup() %>% mutate(genre = avg_rating)
# # now we know what are the mean ratings for each genre now we will combine it to the song id.
# geners_modify_1= geners_modify %>% group_by(genre) %>% mutate(avg_rating = mean(rating)) %>% ungroup() %>% mutate(genre = avg_rating)
# geners_modify_1
# sapply(geners_modify_1, function(data) which(is.na(data)))
# geners_modify_1 = geners_modify_1 %>% group_by(id) %>% mutate(genre = mean(genre))
# geners_modify_1 = geners_modify_1 %>%  select(-avg_rating)
# mean(geners_modify_1$genre[c(1:4)])
# geners_modify_1 %>% group_by(id)
# geners_modify_2= distinct(geners_modify_1, .keep_all = T)
# 
# geners_modify_2 = create_interactions_Variable(geners_modify_2)
# # geners_modify_2 = create_danceability_variance(geners_modify_2)
# # geners_modify_2 = create_temo_variance(geners_modify_2)
# geners_modify_2 = create_song_intensity(geners_modify_2)
# geners_modify_2 = create_acoustic_energy(geners_modify_2)
# geners_modify_2 = create_live_energy(geners_modify_2)
# geners_modify_2 = create_instrumental_energy(geners_modify_2)
# geners_modify_2
# clean.dat <- clean.dat %>% 
#   mutate(the_performer = ifelse(grepl("The", x = clean.dat$performer, ignore.case = TRUE), "YES", "NO"))
# 
# clean.test <- clean.test %>% 
#   mutate(the_performer = ifelse(grepl("The", x = clean.test$performer, ignore.case = TRUE), "YES", "NO"))
songs_train_data = modify_performer_gener(songs_train)
songs_train_data = change_track_duration(songs_train_data)
songs_train_data = create_interactions_Variable(songs_train_data)

songs_train_data = create_song_intensity(songs_train_data)
songs_train_data = create_acoustic_energy(songs_train_data)
songs_train_data = create_live_energy(songs_train_data)
songs_train_data = create_instrumental_energy(songs_train_data)
songs_train_data = songs_train_data %>% select(-id,-performer,-song, -genre)

dim(songs_train_data)
songs_test_data = modify_performer_gener(songs_test)
songs_test_data = change_track_duration(songs_test_data)
songs_test_data = create_interactions_Variable(songs_test_data)

songs_test_data = create_song_intensity(songs_test_data)
songs_test_data = create_acoustic_energy(songs_test_data)
songs_test_data = create_live_energy(songs_test_data)
songs_test_data = create_instrumental_energy(songs_test_data)
songs_test_data = songs_test_data %>% select(-id,-performer,-song, -genre)
dim(songs_test_data)
sapply(songs_train_data, function(dat) which(is.na(dat)))
  
cvforest = randomForest(rating~., 
                        songs_train_data, 
                        mtry = 14, 
                        ntree = 4000)

# pred_train = predict(cvforest)
# rmse_train_cv_forest = sqrt(mean((pred_train - train$Balance)^2)); rmse_train_cv_forest
pred_forest = predict(cvforest, newdata= songs_test_data)
# rmse_cv_forest = sqrt(mean((pred_forest - geners_modify_2$rating)^2)); rmse_cv_forest
submissionFile = data.frame(id = songs_test_2$id, rating = pred_forest)
write.csv(submissionFile, 'Model_4_WW15_2_forest_mtry_8.csv',row.names = F)
  
  
  
  
# pinoy hip hop	71.00000			
# stomp pop	68.66667			
# portland hip hop	68.00000			
# minnesota hip hop	67.50000			
# k-hop	67.00000			
# alt z	65.50000			
# queer country	65.00000			
# italo dance	64.00000			
# middle earth	64.00000			
# corrido	63.00000	
# deep regional mexican	63.00000			
# uk funky	61.33333			
# anarcho-punk	61.00000			
# viral trap	60.25000	
# theme	60.00000			
# bassline	59.00000			
# cartoon	59.00000			
# barbadian pop	58.58621			
# melbourne bounce international	58.00000			
# vapor twitch	58.00000
# etherpop	57.55556			
# celtic	57.00000			
# gauze pop	57.00000			
# german soundtrack	57.00000			
# shiver pop	57.00000			
# dfw rap	56.59259			
# danish pop	56.33333			
# steampunk	56.00000			
# regional mexican pop	55.50000			
# dubstep	55.00000	
# slap house	54.66667			
# swedish hard rock	54.25000			
# deep norteno	54.00000			
# derby indie	54.00000			
# filter house	54.00000			
# norteno-sax	54.00000			
# austropop	53.50000			
# neue deutsche welle	53.50000			
# indie singer-songwriter	53.00000			
# canadian latin	52.80000	
# progressive electro house	52.80000			
# new age	52.50000			
# german pop	52.16667			
# breakbeat	52.00000			
# liquid funk	52.00000			
# scandipop	51.66667			
# k-pop boy group	51.33333			
# alabama rap	51.00000			
# italo house	51.00000			
# jazz trio	51.00000	
# ska revival	51.00000	
```

## Test





```{r}
songs_test = read_csv("data/scoringData.csv")
songs_test = change_track_explicit_to_numeric(songs_test)
songs_test = songs_test %>% separate_rows(genre, sep = ",\\s*")
songs_test = songs_test %>%  mutate(genre = str_replace_all(genre, "\\[|\\]|'|\"", ""))
songs_test = songs_test %>% mutate(track_duration = track_duration/(60*1000))

# songs_test_without_rating = songs_test
genre_mean_rating 
songs_test

# copying the corresponding values of the genre type to songs
index_value = match(songs_test$genre, genre_mean_rating$genre)
# index_value
songs_test$genre = genre_mean_rating$avg_rating[index_value] 
songs_test_1 = songs_test %>% group_by(id) %>% mutate(genre = mean(genre))
songs_test_1
songs_test_2= distinct(songs_test_1, .keep_all = T)
songs_test_2
songs_test_2 = create_interactions_Variable(songs_test_2)
# geners_modify_2 = create_danceability_variance(geners_modify_2)
# geners_modify_2 = create_temo_variance(geners_modify_2)
songs_test_2 = create_song_intensity(songs_test_2)
songs_test_2 = create_acoustic_energy(songs_test_2)
songs_test_2 = create_live_energy(songs_test_2)
songs_test_2 = create_instrumental_energy(songs_test_2)
dim(songs_test_2)

# merged_song_set = merge(songs_test, genre_mean_rating, by.x = "genre", by.y = "genre",all.x = T)
# merged_song_set
# songs_test$genre = merged_song_set$avg_rating
# songs_test
# songs_test_without_rating
#lets verify if the values for that particular song is given right average rating value.
# songs_test = songs_test %>% group_by(genre) %>% mutate(avg_rating = mean(rating)) %>% ungroup() %>% mutate(genre = avg_rating)
geners_modify_2
songs_test_2 = impute_values(songs_test_2)
write.csv(songs_test_2, 'gener_mean_test.csv',row.names = F)


```


# SH Random Forest


```{r}
model_1 = randomForest(rating~., data=geners_modify_2, ntree = 3000)
importance_scores = importance(model_1)
top_predictors = names(importance_scores[order(importance_scores, decreasing = TRUE)[1:5]])
pred = predict(model_1, newdata=songs_test_2)

submissionFile <- data.frame(id = songs_test_2$id,rating=pred)
write.csv(submissionFile, 'Model_4_WW15_2_randomforest_no_tuning.csv',row.names = F)


```

# check for best subset

```{r}
head(geners_modify_2)
geners_modify_2 = geners_modify_2 %>% select(id, performer, song, rating, everything())
trt = designTreatmentsZ(dframe = geners_modify_2,
                        varlist = names(geners_modify_2)[5:length(geners_modify_2)])

newvars = trt$scoreFrame[trt$scoreFrame$code%in% c('clean','lev'),'varName']
newvars
train_input = prepare(treatmentplan = trt, 
                      dframe = geners_modify_2,
                      varRestriction = newvars)
test_input = prepare(treatmentplan = trt, 
                     dframe = songs_test_2,
                     varRestriction = newvars)
head(train_input)

## choosing the hyperparameter



tune_nrounds = xgb.cv(data=as.matrix(train_input), 
                      label = geners_modify_2$rating,
                      nrounds=250,
                      nfold = 5,
                      verbose = 0)

ggplot(data=tune_nrounds$evaluation_log, aes(x=iter, y=test_rmse_mean))+
  geom_point(size=0.4, color='sienna')+
  geom_line(size=0.1, alpha=0.1)+
  theme_bw()
which.min(tune_nrounds$evaluation_log$test_rmse_mean)

xgboost2= xgboost(data=as.matrix(train_input), 
                  label = geners_modify_2$rating,
                  nrounds=13,
                  verbose = 0)
pred = predict(xgboost2, 
               newdata=as.matrix(train_input))
# pred
rmse_xgboost = sqrt(mean((pred - geners_modify_2$rating)^2)); rmse_xgboost
```


# prediction

```{r}
pred = predict(xgboost2, 
               newdata=as.matrix(test_input))

```


# Saving the model

```{r}

submissionFile = data.frame(id = songs_test_2$id, rating = pred)
write.csv(submissionFile, 'Model_4_WW15_2_xgboost.csv',row.names = F)
# head(songs_test)
# head(geners_modify)
# dim(geners_modify)


```


## without optimal round

```{r}
xgboost = xgboost(data=as.matrix(train_input), 
                  label = geners_modify_2$rating,
                  nrounds=10000,
                  verbose = 0,
                  early_stopping_rounds = 200)
xgboost$best_iteration

plot(xgboost$evaluation_log)

pred_train = predict(xgboost, 
               newdata=as.matrix(train_input))
rmse_train_xgboost = sqrt(mean((pred_train - geners_modify_2$rating)^2)); rmse_train_xgboost

## predicting test
pred_test = predict(xgboost, 
               newdata=as.matrix(test_input))
submissionFile = data.frame(id = songs_test_2$id, rating = pred_test)
write.csv(submissionFile, 'Model_4_WW15_2_xgboost_without_optimal.csv',row.names = F)
```


## Tuned Forest

```{r}
library(randomForest)
trControl = trainControl(method = 'cv', number = 5)
head(geners_modify_2)
train = geners_modify_2 %>% select(everything())
head(train)
train$new_rating = train$rating
names(train)[names(train) == "new_rating"] = "rating"
tuneGrid = expand.grid(mtry = 1:20)
tuneGrid
set.seed(1031)
forest_cv = train(rating~., 
                  data = train, 
                  method = 'rf', 
                  trControl = trControl, 
                  tuneGrid = tuneGrid, 
                  ntree = 2000)
forest_cv$bestTune$mtry

cvforest = randomForest(rating~., 
                        geners_modify_2, 
                        mtry = 8, 
                        ntree = 4000)

# pred_train = predict(cvforest)
# rmse_train_cv_forest = sqrt(mean((pred_train - train$Balance)^2)); rmse_train_cv_forest
pred_forest = predict(cvforest, newdata= songs_test_2)
# rmse_cv_forest = sqrt(mean((pred_forest - geners_modify_2$rating)^2)); rmse_cv_forest
submissionFile = data.frame(id = songs_test_2$id, rating = pred_forest)
write.csv(submissionFile, 'Model_4_WW15_2_forest_mtry_8.csv',row.names = F)


```

## Tuned Ranger

```{r}
head(geners_modify_2)
train = geners_modify_2[,-1]
train = train[,3:length(train)]
train$new_rating = train$rating
train = train[,2:length(train)]
names(train)[names(train) == "new_rating"] = "rating"
head(train)
trControl=trainControl(method="cv",number=5)
tuneGrid = expand.grid(mtry=1:ncol(train)-1, 
                       splitrule = c('variance','extratrees','maxstat'), 
                       min.node.size = c(2,5,10,15,20,25))
set.seed(1031)
cvModel = train(rating~.,
                data=train,
                method="ranger",
                num.trees=1000,
                trControl=trControl,
                tuneGrid=tuneGrid)
cvModel$bestTune

##Predicting

cv_forest_ranger = ranger(rating~.,
                          data=train,
                          num.trees = 1000, 
                          mtry=cvModel$bestTune$mtry, 
                          min.node.size = cvModel$bestTune$min.node.size, 
                          splitrule = cvModel$bestTune$splitrule)

pred_train = predict(cv_forest_ranger, data = train, num.trees = 1000)
rmse_train_cv_forest_ranger = sqrt(mean((pred_train$predictions - train$Balance)^2)); rmse_train_cv_forest_ranger

```