---
title: "Model_6"
author: "Srivastav Budugutta"
date: "2023-03-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Overview
>In this we will be investigating on the performers data.
Different types of performers are: individual names, performers with "featuring in between performers, & is there between performers, "+" is there when the artisits combine with other performers.


#initiating library

```{r}
rm(list = ls())
library(tidyverse)
library(mice)
```
## Function definitions
```{r}
clean_rating = function(dat){
  dat = dat %>% mutate(rating = if_else(rating<0,0,rating))
  
  return(dat)
}

change_track_explicit_to_numeric = function(dat){
  dat = dat %>% mutate(track_explicit = ifelse(track_explicit == FALSE,0,1))
  return(dat)
}

create_interactions_Variable = function(dat){
  dat = dat %>% mutate(interactive = danceability*energy)
  return(dat)
}

create_song_intensity = function(dat){
  dat = dat %>% mutate(song_intensity = danceability * energy * (10^(loudness/20)))
  return(dat)
}

create_acoustic_energy = function(dat){
  dat = dat %>% mutate(acoustic_energy = (1 - instrumentalness) * energy)
  return(dat)
}

create_instrumental_energy = function(dat){
  dat = dat %>% mutate(instrumental_energy = energy - (energy*(1-instrumentalness)))
  return(dat)
}

create_live_energy = function(dat){
  dat = dat %>% mutate(liveenergy = liveness*energy)
  return(dat)
}

# not working
create_danceability_variance = function(dat){
  dat = dat %>% mutate(danceability_variance = sd(danceability))
  return(dat)
}

# not working
create_temo_variance = function(dat){
  dat = dat %>% mutate(tempo_variance = sd(tempo))
  return(dat)
}



```

# Train Data Set
```{r}
songs = read_csv("data/analysisData.csv")


dim(songs)
songs = clean_rating(songs)
songs = change_track_explicit_to_numeric(songs)
songs %>% group_by(performer) %>% summarise(mean_rating = mean(rating))


# songs$track_explicit
head(songs)
min(songs$rating)
# geners_modify = songs %>% mutate(genre = str_extract_all(genre, regex("\'(.*?)\'", ignore_case = T)))
# geners_modify$genre

# we are trying to separate all the genre types
geners_modify = songs %>% separate_rows(genre, sep = ",\\s*")
geners_modify = geners_modify %>%  mutate(genre = str_replace_all(genre, "\\[|\\]|'|\"", ""))

genre_mean_rating = geners_modify %>% group_by(genre) %>% summarize(avg_rating = mean(rating))
genre_mean_rating %>% filter(genre == "acoustic punk")
genre_mean_rating[geners_modify$genre,]
head(geners_modify)
geners_modify = geners_modify %>% mutate(track_duration = track_duration/(60*1000))

(geners_modify %>% group_by(genre) %>% summarize(avg_rating = mean(rating)) %>% arrange(avg_rating))
# geners_modify_1= geners_modify %>% group_by(genre) %>% summarize(avg_rating = mean(rating)) %>% ungroup() %>% mutate(genre = avg_rating)
# now we know what are the mean ratings for each genre now we will combine it to the song id.
geners_modify_1= geners_modify %>% group_by(genre) %>% mutate(avg_rating = mean(rating)) %>% ungroup() %>% mutate(genre = avg_rating)
geners_modify_1 = geners_modify_1 %>% group_by(id) %>% mutate(genre = mean(genre))
geners_modify_1 = geners_modify_1 %>%  select(-avg_rating)
mean(geners_modify_1$genre[c(1:4)])
geners_modify_1 %>% group_by(id)
geners_modify_2= distinct(geners_modify_1, .keep_all = T)

geners_modify_2 = create_interactions_Variable(geners_modify_2)
# geners_modify_2 = create_danceability_variance(geners_modify_2)
# geners_modify_2 = create_temo_variance(geners_modify_2)
geners_modify_2 = create_song_intensity(geners_modify_2)
geners_modify_2 = create_acoustic_energy(geners_modify_2)
geners_modify_2 = create_live_energy(geners_modify_2)
geners_modify_2 = create_instrumental_energy(geners_modify_2)
geners_modify_2

# now in model5 we will work on artists names
geners_modify_2 


performer_average = geners_modify_2 %>% mutate(performer = mean(rating))
performer_rating = geners_modify_2 %>% group_by(performer) %>% summarize(avg_rating = mean(rating))
performer_rating

```

## Test





```{r}
songs_test = read_csv("data/scoringData.csv")
songs_test = change_track_explicit_to_numeric(songs_test)
songs_test = songs_test %>% separate_rows(genre, sep = ",\\s*")
songs_test = songs_test %>%  mutate(genre = str_replace_all(genre, "\\[|\\]|'|\"", ""))
songs_test = songs_test %>% mutate(track_duration = track_duration/(60*1000))

# songs_test_without_rating = songs_test
genre_mean_rating 
songs_test

# copying the corresponding values of the genre type to songs
index_value = match(songs_test$genre, genre_mean_rating$genre)
# index_value
songs_test$genre = genre_mean_rating$avg_rating[index_value] 
songs_test_1 = songs_test %>% group_by(id) %>% mutate(genre = mean(genre))
songs_test_1
songs_test_2= distinct(songs_test_1, .keep_all = T)
songs_test_2
songs_test_2 = create_interactions_Variable(songs_test_2)
# geners_modify_2 = create_danceability_variance(geners_modify_2)
# geners_modify_2 = create_temo_variance(geners_modify_2)
songs_test_2 = create_song_intensity(songs_test_2)
songs_test_2 = create_acoustic_energy(songs_test_2)
songs_test_2 = create_live_energy(songs_test_2)
songs_test_2 = create_instrumental_energy(songs_test_2)
dim(songs_test_2)
songs_test_2
index_value = match(songs_test_2$performer, performer_rating$performer)
index_value
# index_value
songs_test_2$performer = performer_rating$avg_rating[index_value]
songs_test_2[!complete.cases(songs_test_2), ]
summary(songs_test_2)
songs_test_with_mice = mice(songs_test_2, m = 5, method = "pmm", seed = 123)
songs_test_2 = complete(songs_test_with_mice,2)

songs_test_2


```


# check for best subset

```{r}
# here we use hybrid model

start_mod = lm(rating~1,data=performer_average)
empty_mod = lm(rating~1,data=performer_average)

full_mod = lm(rating~tempo+time_signature+mode+speechiness+ track_explicit+genre+track_duration+danceability+energy+key+loudness+ interactive+song_intensity+acoustic_energy+instrumental_energy+liveenergy+performer,performer_average)



hybridStepwise = step(start_mod,
                      scope=list(upper=full_mod,lower=empty_mod),
                      direction='both')
summary(hybridStepwise)

hybridStepwise$anova
```


# prediction

```{r}
#good subset:
# performer
# danceability
# energy
# speechiness
# loudness
# track_explicit
# track_duration


model_1 = lm(rating~performer+liveenergy+key+mode,performer_average)
model_1_predict = predict(model_1,songs_test_2)
# length(songs_test_2$id)
# length(model_1_predict)
# songs_test_2[2812,]

# songs_test_2[,6]



```


# Saving the model

```{r}

submissionFile = data.frame(id = songs_test_2$id, rating = model_1_predict)
submissionFile[!complete(submissionFile),]
write.csv(submissionFile, 'model_6.csv',row.names = F)
# head(songs_test)
# head(geners_modify)
# dim(geners_modify)


```




```{r}



```




```{r}



```